---
title: "A4-Q2"
author: "Zhaoqi Li"
date: "2023/3/28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1.

```{r}
x <- c(1.6, 0.1, 0.3, 0.0, -1.1, -0.1, -0.4, -0.4)
acf(x, plot = FALSE , type = "covariance", lag.max = 2)
```
The result of Q2(1) is same as Q1(1)

### 2. 

```{r}
x.yw <- ar.yw(x, aic= FALSE, order=1)
x.yw$ar
x.yw$var.pred
```
The coefficient estimates and  estimated error variance are similar but little different of the estimated error variance of Q1(2).

### 3.

```{r}
x.yw$ar[1]+c(-1.96,1.96)*sqrt(x.yw$asy.var.coef[1, 1])
```
The 95% CI are also similar but but little different of Q1(3), because the estimated error variance are little different.

### 4. 

```{r}
x.yw$resid
```
The result of Q2(4) is same as Q1(4)

### 5.

```{r}
Box.test(x.yw$resid, lag = 2, type = c("Ljung-Box"), fitdf = 1)
```
The p-value is 0.2959 which is greater than 0.05, so we fail to reject H0. The result of Q2(5) is same as Q1(5)

### 6 & 7

```{r}
library("forecast")
forecast(x.yw, level= 0.95, h=2)
```
The result of 6 and 7 are similar to that of Q1.
