---
title: "A4-Q3"
author: "Zhaoqi Li"
date: "2023/4/9"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Step 1: Load the data and plot transformation

```{r}
library("astsa") 
plot(UnempRate)
```
The series is not stationary because it shows a seasonal pattern. The trend is  slightly upward.

### Step 2: Check how many differences are required to make the series stationary

```{r}
library("forecast")
ndiffs(UnempRate)

```

### Step 3: Take the first seasonal difference series and examine the plots of the series, ACF, and PACF

```{r}
library("tidyverse")
UnempRate %>% diff(lag=12) %>% ggtsdisplay()
```
According to the plot, the ACF decrease, so we need non-seasonal difference.

### Step 4: Take another first seasonal difference series and examine the plots of the series, ACF, and PACF
```{r}
UnempRate %>% diff(lag=12) %>% diff() %>%
ggtsdisplay()
```
According to the plots, there are two spikes at lags 1, 2, 3 in PACF, so we suggest a non-seasonal AR(3). According to the ACF plot, there are significant spikes at lag 1, 2, 3, 4, 5 and lag 12 , so we suggest a non-seasonal MA(5) and a seasonal SMA(1). Thus, our tentative model is $$ ARIMA(3,1,5)Ö(0,1,1)_{12}$$
```{r}
UnempRate %>% 
  Arima(order=c(3, 1, 5), seasonal = c(0, 1, 1)) %>%
  residuals() %>%
  ggtsdisplay()
```
```{r}
fit <- Arima(UnempRate, order=c(3, 1, 5), seasonal = c(0, 1, 1)) 
fit$aic
fit$aicc
fit$bic
checkresiduals(fit)
```
According to the value of the BIC, we choose this model to fit the log series. According to the diagnostic plots of the residual and the Ljung-Box portmanteau test statistic, the p-value is 0.2621 ,so we indicate the residuals are white noise but the  fitted model is not very good because the p-value is not very samll.

### Step 5: Forecast the next 12 months

```{r}
forecast(fit, h=12)
fit %>%
  forecast(h=12)%>%
  autoplot()
```
Compared with the perivous model, the new model with the function auto.arima() is better because the vaule of AIC, AICC, and BIC is samller.

```{r}
fit.12 <- auto.arima(UnempRate)
fit.12
```
```{r}
checkresiduals(fit.12)
```
According to the plots, both models are available. 
```{r}
fit.12 %>%
  forecast(h=12)%>%
  autoplot()
```

