---
title: "Final Project"
author: "Zhaoqi Li"
date: "2022/12/20"
output: pdf_document
---

```{r}
library(openintro)
library(tidyverse)
library(gridExtra)
library(car)
```

# Clean data and remove missing value

```{r}
# Data clean
my_data <- read_csv("MY2022 Fuel Consumption Ratings.csv")
fuel_cons <-rename(my_data, engine_size = 'Engine Size(L)'
                   , fuel_type = 'Fuel Type'
                   , fuel_consumption_comb = 'Fuel Consumption(Comb (L/100 km))'
                   , co2_emissions_gkm = 'CO2 Emissions(g/km)'
                   , fuel_consumption_hwy = 'Fuel Consumption(Hwy (L/100 km))'
                   , fuel_consumption_city = 'Fuel Consumption (City (L/100 km)'
                   , co2_rating = 'CO2 Rating'
                   , fuel_consumption_mpg = 'Fuel Consumption(Comb (mpg))'
                   , model_year = 'Model Year'
                   , smog_rating = 'Smog Rating') %>%
  mutate(co2_emissions = co2_emissions_gkm*100,
         transmission = case_when(Transmission == "A10" ~ "A",
                                  Transmission == "A9" ~"A",
                                  Transmission == "A7" ~"A",
                                  Transmission == "A6" ~"A",
                                  Transmission == "AM6" ~"AM",
                                  Transmission == "AM7" ~"AM",
                                  Transmission == "AM8" ~"AM",
                                  Transmission == "AS10" ~"AS",
                                  Transmission == "AS9" ~"AS",
                                  Transmission == "AS8" ~"AS",
                                  Transmission == "AS7" ~"AS",
                                  Transmission == "AS6" ~"AS",
                                  Transmission == "AS5" ~"AS",
                                  Transmission == "AV1" ~"AV",
                                  Transmission == "AV10" ~"AV",
                                  Transmission == "AV8" ~"AV",
                                  Transmission == "AV7" ~"AV",
                                  Transmission == "AV6" ~"AV",
                                  Transmission == "M5" ~"M",
                                  Transmission == "M6" ~"M",
                                  Transmission == "M7" ~"M")) %>%
  drop_na()

```
## Step 0: Divide data into training and testing

```{r}
# Create training and test set 
set.seed(147)

# Count the number of observations in the data
n <- nrow(fuel_cons)

# Randomly choose 80% as training and round number
training <- sample(1:n, size = round(0.8*n))

# Create a training set
train <- fuel_cons[training,]

# Create a testing set
test <- fuel_cons[-training,]

```

## Description of Important Variables
```{r}
# Table of the important variables

Variable <-(c("Fuel Consumption","Smog Rating","Transmission", "Cylinders","CO2 Rating","Fuel Type "))
Type <-(c("Numerical Variable","Numerical Variable","Categorical Variable","Numerical Variable","Numerical Variable","Categorical Variable"))
Description <-(c("The combine fuel consumption (city 55%, hwy 45%) in liters per 100 kilometers(L/100km)","The emissions of smog-forming pollutants rated on a scale from 1 (worst) to 10 (best)","The type of transmissions.A = automatic; M = manual; AM = automated manual; AS = automatic with select shift; AV = continuously","The size of engine (L)","The emissions of CO2 pollutants rated on a scale from 1 (worst) to 10 (best)","The fuel type of vehicle. X = regular gasoline; Z = premium gasoline; D = diesel ; E = ethanol (E85);N = natural gas"))
knitr::kable(tibble(Variable,Type, Description),caption="The Description of Important Variables")
```

## Exploratory Data Analysis of training set

```{r}
# Box plot and bar plot of predictors in training dataset
bar_transmission <- train %>%
  ggplot(aes(x=transmission)) +
  geom_bar(color='black', fill='pink') +
  labs(title="Transmission") +
  coord_flip()

bar_fuel_type <- train %>%
  ggplot(aes(x=fuel_type)) +
  geom_bar(color='black', fill='pink') +
  labs(title="Fuel Type") +
  coord_flip()
```
```{r}
box_fuel_cons <- train %>%
  ggplot(aes(y=fuel_consumption_comb)) +
  geom_boxplot(color= "black", fill="pink") +
  labs(title="Fuel Consumption")
box_smog_rating <- train %>%
  ggplot(aes(y=smog_rating)) +
  geom_boxplot(color="black", fill="pink") +
  labs(title="Smog rating")

box_co2_rating <- train %>%
  ggplot(aes(y=co2_rating)) +
  geom_boxplot(color="black", fill="pink") +
  labs(title="CO2 rating")

box_cylinders <- train %>%
  ggplot(aes(y=Cylinders)) +
  geom_boxplot(color="black", fill="pink") +
  labs(title="Cylinders")

grid.arrange(box_fuel_cons,box_smog_rating,box_co2_rating,box_cylinders,bar_fuel_type,bar_transmission,nrow=2)
```

## Step 1: Choose a starting model

```{r}
# Full model with all possible predictors
model_full <- lm(fuel_consumption_comb ~ co2_emissions + smog_rating + transmission + engine_size + Cylinders + co2_rating + fuel_type, data=train)
```


## Step 2: Check condition 1&2 and assumptions

```{r}
## Condition 1: draw a scatter plot between yi and y_hat
y_hat <- fitted(model_full)
yi <- train$fuel_consumption_comb
plot(yi,y_hat)
abline(a = 0, b = 1)
lines(lowess(yi ~ y_hat), lty=2)
```
```{r}
# Condition 2: draw scatter plots between predictors (numerical predictors)
pairs(~co2_emissions + smog_rating + engine_size + Cylinders + co2_rating, data=train)
```
```{r}
## Residual vs. Fitted model
res <- model_full$residuals
y_hat <- fitted(model_full)
plot(y_hat, res)
```

```{r}
## Residual vs. Predictors
par(mfrow = c(1, 2))
plot(train$co2_rating, res)
plot(train$smog_rating,res)
plot(train$Cylinders,res)
plot(train$engine_size,res)
plot(train$co2_emissions,res)
```

```{r}
qqnorm(res)
qqline(res)
```
## Step 3 : Model transformations to correct assumption violations
```{r}
# Summary box-cox transformation for numerical variables
summary(powerTransform(cbind(train$fuel_consumption_comb,
                             train$engine_size,
                             train$Cylinders,
                             train$co2_emissions,
                             train$co2_rating,
                             train$smog_rating)))

```
```{r}
# Create transformed variables
train_trans <- train %>%
  mutate(trans_fuel_consumption_comb = fuel_consumption_comb^(-0.5),
         trans_engine_size =log(engine_size),
         trans_Cylinders = Cylinders^(-0.5),
         trans_co2_emissions = co2_emissions^(-0.33),
         trans_smog_rating = smog_rating^(1.7))
         
```

```{r}
# Fit a new model with the transformed variables
model_full_trans <- lm(trans_fuel_consumption_comb ~trans_co2_emissions + trans_smog_rating + transmission + trans_engine_size + trans_Cylinders + co2_rating + fuel_type, data = train_trans)
```

## Step 4: Ensure no multicollinearity is present in the model

```{r}
# Check the VIF of the model and remove the predictors with high VIF
vif(model_full_trans)
```
```{r}
# Remove transformed CO2 emissions because it has a high VIF
model_full_trans1 <- lm(trans_fuel_consumption_comb ~ trans_smog_rating + transmission + trans_engine_size + trans_Cylinders + co2_rating + fuel_type, data = train_trans)

vif(model_full_trans1)
```

```{r}
# Remove transformed engine size  because it has a high VIF
model_full_trans2 <- lm(trans_fuel_consumption_comb ~ trans_smog_rating + transmission + trans_Cylinders  + co2_rating + fuel_type, data = train_trans)

vif(model_full_trans2)
```
# Step 5: Model selection
```{r}
# Manually select the predictors with low P-value as reduced model
summary(model_full_trans2)
```
```{r}
# Remove the predictors transmission and cylinders becuase they are not significant
model_reduced <- lm(trans_fuel_consumption_comb ~ trans_smog_rating + co2_rating + fuel_type, data = train_trans)
```


```{r}
# Compare the F test of full model and reduced model
anova(model_reduced, model_full_trans2)
```
```{r}
# Add back a removed predictor to ensure F test is large than 0.05
model_reduced2 <- lm(trans_fuel_consumption_comb ~ trans_smog_rating + co2_rating + fuel_type + transmission, data = train_trans)
anova(model_reduced2, model_full_trans2)
```
## Check the Condition 1&2 and assumptions of full model 

```{r}
## Condition 1: draw a scatter plot between yi and y_hat
y_hat <- fitted(model_full_trans2)
yi <- train_trans$trans_fuel_consumption_comb
plot(yi,y_hat)
abline(a = 0, b = 1)
lines(lowess(yi ~ y_hat), lty=2)
```
```{r}
# Condition 2: draw scatter plots between predictors (numerical predictors)
pairs(~trans_smog_rating+co2_rating+trans_Cylinders, data=train_trans)
```
```{r}
## Residual vs. Fitted model 
res <- model_full_trans2$residuals
y_hat <- fitted(model_full_trans2)
plot(y_hat, res)
```

```{r}
## Residual vs. Predictors
plot(train_trans$co2_rating, res)
plot(train_trans$trans_smog_rating,res)
plot(train_trans$trans_Cylinders,res)
```
```{r}
# Use normal QQ plot check normality
qqnorm(res)
qqline(res)
```
## Check the Condition 1&2 and assumptions of reduced model 
```{r}
## Condition 1: draw a scatter plot between yi and y_hat
y_hat <- fitted(model_reduced2)
yi <- train_trans$trans_fuel_consumption_comb
plot(yi,y_hat)
abline(a = 0, b = 1)
lines(lowess(yi ~ y_hat), lty=2)
```


```{r}
# Condition 2: draw scatterplots between predictors (can only be done for numerical predictors)
pairs(~trans_smog_rating+co2_rating, data=train_trans)
```
```{r}
## Residual vs. Fitted
res1 <- model_reduced2$residuals
y_hat <- fitted(model_reduced2)
plot(y_hat, res1)
```

```{r}
## Residual vs. Predictors
par(mfrow = c(1, 2))
plot(train_trans$co2_rating, res1)
plot(train_trans$trans_smog_rating,res1)
```
```{r}
# Use normal QQ plot check normality
qqnorm(res1)
qqline(res1)
```
## Compare the R^2, AIC, and BIC
```{r}
summary(model_full_trans2)$adj.r.squared
summary(model_reduced2)$adj.r.squared
```

```{r}
AIC(model_full_trans2)
AIC(model_reduced2)
```


```{r}
BIC(model_full_trans2)
BIC(model_reduced2)
```

## Step 6:Leverage Point, Outlier, and Influential Point
```{r}
# Leverage Point
h <- hatvalues(model_full_trans2)
Hcut <- 2*(length(model_full_trans2$coefficients)/nrow(train_trans))
which(h > Hcut)
```
```{r}
 # Outlier
r <- rstandard(model_full_trans2)
which(r < -4 | r > 4)
```

```{r}
# Cooks's Distance
D <- cooks.distance(model_full_trans2)
Dcut <- qf(0.5,
               length(model_full_trans2$coefficients),
           nrow(train_trans)-length(model_full_trans2$coefficients))
which(D > Dcut)
```


```{r}
# Leverage Point
h <- hatvalues(model_reduced2)
Hcut <- 2*(length(model_reduced$coefficients)/nrow(train_trans))
which(h > Hcut)
```

```{r}
 # Outlier
r <- rstandard(model_reduced2)
which(r < -4 | r > 4)
```
```{r}
# Cooks's Distance
D <- cooks.distance(model_reduced2)
Dcut <- qf(0.5,
               length(model_reduced2$coefficients),
           nrow(train_trans)-length(model_reduced2$coefficients))
which(D > Dcut)
```

## Step 7: Model Validation

### Compare EDA 
```{r}
bar_transmission <- test %>%
  ggplot(aes(x=transmission)) +
  geom_bar(color='black', fill='pink') +
  labs(title="Transmission") +
  coord_flip()

bar_fuel_type <- test %>%
  ggplot(aes(x=fuel_type)) +
  geom_bar(color='black', fill='pink') +
  labs(title="Fuel Type") +
  coord_flip()
box_fuel_cons <- test %>%
  ggplot(aes(y=fuel_consumption_comb)) +
  geom_boxplot(color= "black", fill="pink") +
  labs(title="Fuel Consumption")
box_smog_rating <- test %>%
  ggplot(aes(y=smog_rating)) +
  geom_boxplot(color="black", fill="pink") +
  labs(title="Smog rating")

box_co2_rating <- test %>%
  ggplot(aes(y=co2_rating)) +
  geom_boxplot(color="black", fill="pink") +
  labs(title="CO2 rating")

box_cylinders <- test %>%
  ggplot(aes(y=Cylinders)) +
  geom_boxplot(color="black", fill="pink") +
  labs(title="Cylinders")

grid.arrange(box_fuel_cons,box_smog_rating,box_co2_rating,box_cylinders,bar_fuel_type,bar_transmission,nrow=2)
```

###Apply the same transformation on testing data

```{r}
test_trans <- test %>%
  mutate(trans_fuel_consumption_comb = fuel_consumption_comb^(-0.5),
         trans_engine_size =log(engine_size),
         trans_Cylinders = Cylinders^(-0.5),
         trans_co2_emissions = co2_emissions^(-0.33),
         trans_smog_rating = smog_rating^(1.7))
```

###Refit the model on testing dat
```{r}
model_full_test <- lm(trans_fuel_consumption_comb ~ trans_smog_rating + transmission + trans_Cylinders+co2_rating + fuel_type, data = test_trans)
```

### Compare the coefficients of training model and testing model
```{r}
summary(model_full_trans2)
```
```{r}
summary(model_full_test)
```

```{r}
Variable <-(c("Intercept","Smog Rating","TransmissionAM","TransmissionAS","TransmissionAV","TransmissionM","Cylinders","CO2 Rating","Fuel Type E","Fuel Type X","Fuel Type Z"))
Coefficient_Estimate_Train <-(c("0.2049","0.0001533","0.005526","0.002666","0.006973","0.001748","0.01616","0.02456","-0.06434","-0.01947","-0.02264"))
Coefficient_Estimate_Test <-(c("0.2096897","0.0002793","0.0084263","0.0037115","0.0057024","-0.0013993","-0.0220062","0.0265233","-0.0587159","-0.0189943","-0.0226666"))
knitr::kable(tibble(Variable,Coefficient_Estimate_Train,Coefficient_Estimate_Test),caption="The Summary of Coefficient Estimate for Train and Test")
```

## Check condition 1&2 and assumptions of test model

```{r}
# Condition 1
y_hat <- fitted(model_full_test)
yi <- test_trans$trans_fuel_consumption_comb
plot(yi,y_hat)
abline(a = 0, b = 1)
lines(lowess(yi ~ y_hat), lty=2)
```
```{r}
# Condition 2: draw scatt erplots between predictors (numerical predictors)
pairs(~trans_smog_rating+co2_rating+trans_Cylinders, data=test_trans)
```
```{r}
## Residual vs. Fitted
res <- model_full_test$residuals
y_hat <- fitted(model_full_test)
plot(y_hat, res)
```
```{r}
## Residual vs. Predictors
plot(test_trans$co2_rating, res)
plot(test_trans$trans_smog_rating,res)
plot(test_trans$trans_Cylinders)
```

```{r}
# Use normal QQ plot check normality
qqnorm(res)
qqline(res)
```

 