---
title: "HW1-Q5"
author: "Zhaoqi Li"
date: "2023/9/30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ISLR)
```

### a.
```{r}
model1 <- lm(Sales ~ Price + Urban + US, data = Carseats)
```

### b.
```{r}
summary(model1)
```

### Interpretation
Price: If other variables are fixed (Urban and USYes), when the price of carseats increase 1 dollar, the sales of car seats will decrease 54.46 on average.

UrbanYes: If other variables are fixed (Price and USYes), when the location is in urban, the sales of car seats will decrease 21.91 on average than that is not in urban.

USYes: If other variables are fexed (Price and UrbanYes), when the location is in US, the sales of car seats will increase 1201 on average than that is not in US.

### c.

$$ Sales_i = 13.04 - 0.054\times Price_i - 0.021\times Urban_i + 1.2\times US_i $$
$Urban_i$: If the location is in urban i=1; if the location is not in urban i=0.

$US_i$:If the location is in US i=1; if the location is not in US i=0.

### d.
According to the output, the p value of Price and USYes are significant at 5% level, for these two predictor we can reject H0.

### e. 

```{r}
model2 <- lm(Sales ~ Price + US, data = Carseats)
```

### f. 
```{r}
summary(model1)$r.squared
summary(model2)$r.squared
```

The $R^2$ of these two models are similar and that of model 1 is a little large.
No, we should also consider adjusted R-square which consider the complexity of the model. For the R-square, it determines the proportion of variance in the dependent variable that can be explained by the independent variable. Therefore, the increase of covariates will have larger R-square
### g.
```{r}
confint(model2)
```
### h.
```{r}
model3 <- lm(data=Carseats, Sales ~ Price*US)
summary(model3)
```
### Interpretation
Price: When the price of carseats increase 1 dollar, the sales of car seats will decrease 53.99 on average for a non-US store.

USYes: When the price is zeor, the sales of car seats for US store is 1295 more on average than that for a non-US store.

Price:USYes: When the price of carseats increase 1 dollar for US stores in US,the sales of car seat will 0.84 fewer than than that for a non-US store. 
